---
title: "stats final"
author: "Victoria Cutler"
date: "2022-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Relevant Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(stringr)
library(spData)
library(tmap)
library(broom)

library(tidyverse)
library(readr)
library(gt)
library(openintro)
library(ggplot2)
library(modelr)
library(knitr)
library(xtable)
```

## Reading In The Data

```{r}
# solar data is from: https://sunroof.withgoogle.com/data-explorer/featured/1/oklahoma-city
rawdata_solar <- read_csv(here("data", "project-sunroof-census_tract.csv")) 
solar_meta <- read_csv(here("data", "metadata.csv"))

# income data is from: https://www.ffiec.gov/censusapp.htm
rawdata_income <- read_csv(here("data", "2022CensusMsaStateCountyTractWithIncomeInd.csv"))
```

## Data Quality Exploration

```{r}
summary(rawdata_solar) # checking mins and maxes and signs. percent_covered has a weird range.
summary(rawdata_income) # some tract med family income is 0, so removing 0s where there are tracts, incomes, or income level inds = 0 for data quality
head(rawdata_income)
```

## Data Clean Up & Tidying

```{r}
# solar data clean up and tidying
solardata <- rawdata_solar |> 
  select(region_name, state_name, count_qualified, percent_qualified, number_of_panels_median, number_of_panels_total, yearly_sunlight_kwh_median, yearly_sunlight_kwh_total, existing_installs_count) |>  # limit to variables of interest, removing percent covered since range is above 100%
  drop_na() |> # remove any rows with NA values
  mutate(pct_qual_without_solar = (existing_installs_count/count_qualified) * 100) |> # small numbers mean that this census tract is not living up to it's potential in terms of solar! 
  rename(census_tract_id = region_name) #|> 
  #mutate(log_normalized_percentage = log(percent_without_solar))#|> 
  
# income data clean up and tidying
incomedata <- rawdata_income |> 
  clean_names() |> 
  mutate(state_code = as.character(state_code)) |> 
  mutate(county_code = as.character(county_code)) |> 
  mutate(tract = as.character(tract)) |> 
  mutate(state_code = str_pad(state_code, 2, pad = "0")) |> 
  mutate(county_code = str_pad(county_code, 3, pad = "0")) |> 
  mutate(tract = str_pad(tract, 6, pad = "0")) |> 
  mutate(census_tract_id = paste0(state_code, county_code, tract)) |> 
  filter(income_percent != 0 | tract_med_family_income != 0 | income_percent != 0) |> # removing 0s since this is inaccurate data
  select(year, state_code, tract_med_family_income, income_percent, income_level_ind, census_tract_id) |> 
  drop_na() |> 
  filter(state_code != 72)
```

## Exploration of Distribution/Normality of Solar Data

```{r}
# data distribution
ggplot(data = solardata, aes(x = pct_qual_without_solar)) +
  geom_histogram() # essentially no left tail, large right skew, good candidate for a log transformation to get data normal
  
  # the distribution looks more normal when we log it:
ggplot(data = solardata, aes(x = log(pct_qual_without_solar))) +
  geom_histogram() 

  # looking to confirm or deny log-normality with a qq plot. Since the beginning and ending quartiles diverge from the values we'd expect if they were normal, this signals that our beginning quantiles have higher than expected values and our end quartiles have lower than expected values 
ggplot(data = solardata, aes(sample = log(pct_qual_without_solar))) +
  geom_qq() +
  geom_qq_line()
```

## Joining Clean DataFrames of Interest

```{r}
cleandata <- inner_join(incomedata, solardata, by = "census_tract_id") |> 
  mutate(state_name = as.factor(state_name)) |> 
  mutate(income_level_ind = as.factor(income_level_ind)) |> 
  mutate(log_pct_qual_without_solar = log(pct_qual_without_solar))

cleandata_inflogrm <- cleandata

cleandata_inflogrm[is.na(cleandata_inflogrm) | cleandata_inflogrm == "Inf" | cleandata_inflogrm == "-Inf"] <- NA

cleandata_inflogrm <- cleandata_inflogrm |> drop_na()

# basic data visuals
  # income median
ggplot(data = cleandata_inflogrm, aes(x = tract_med_family_income, y = log_pct_qual_without_solar)) +
  geom_point() +
  labs(x = "Median Family Income",
       y = "Percent of Qualified Buildings Without Rooftop Solar")

# median solar energy potential
ggplot(data = cleandata_inflogrm, aes(x = yearly_sunlight_kwh_median, y = log_pct_qual_without_solar)) +
  geom_jitter(width = 0.5, height = 0, alpha = .8)  +
  labs(x = "Median Yearly Solar Generation Potential (kwh)",
       y = "Percent of Qualified Buildings Without Rooftop Solar")

  # income buckets
ggplot(data = cleandata_inflogrm, aes(x = income_level_ind, y = log_pct_qual_without_solar)) +
  geom_boxplot() +
  labs(x = "Income Level Bin", y = "Percent of Qualified Buildings Without Rooftop Solar")

  # state buckets
ggplot(data = cleandata_inflogrm, aes(x = state_name, y = pct_qual_without_solar)) +
  geom_boxplot() +
  labs(x = "State", y = "Percent of Qualified Buildings Without Rooftop Solar") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

make a linear regression: - also want to look at census tract safety and/or census tract race and/or census tract homeownership

```{r}
mod <- lm(log_pct_qual_without_solar ~ tract_med_family_income + yearly_sunlight_kwh_median + tract_med_family_income:state_code + state_code,
          data = cleandata_inflogrm) # interesting that using total sunlight predicts more than using median sunlight
summary(mod)

# checking to see if residuals are normally distributed
res <- resid(mod)
#qqnorm(res)

#plot(mod, which=2)

ggplot(data = mod, aes(sample = res)) +
  geom_qq() +
  geom_qq_line()

```

# just looking at california:

```{r}
ca_data_inflogrm <- cleandata_inflogrm |> 
  filter(state_name == "California")

ggplot(data = ca_data_inflogrm, aes(x = pct_qual_without_solar)) +
  geom_histogram() 

ggplot(data = ca_data_inflogrm, aes(x = log_pct_qual_without_solar)) +
  geom_histogram() 

ggplot(data = ca_data_inflogrm, aes(sample = log_pct_qual_without_solar)) +
  geom_qq() +
  geom_qq_line()

mod_ca <- lm(log_pct_qual_without_solar ~ tract_med_family_income + yearly_sunlight_kwh_median, data = ca_data_inflogrm)
summary(mod_ca)

# residuals are normally distributed
res_ca <- resid(mod_ca)

ggplot(data = mod_ca, aes(sample = res_ca)) +
  geom_qq() +
  geom_qq_line()

#mod_heart <- glm(pct_qual_without_solar ~ income_level_ind + yearly_sunlight_kwh_median, data = ca_data, family= gaussian(link='log'))
#summary(mod_heart)

summary(mod_ca)

ggplot(data = ca_data_inflogrm, aes(x = tract_med_family_income,y = log_pct_qual_without_solar, color = yearly_sunlight_kwh_median)) +
  geom_point() +
  geom_line(data = augment(mod_ca), aes(y = .fitted, color = yearly_sunlight_kwh_median))

ggplot(data = ca_data_inflogrm, aes(x = yearly_sunlight_kwh_median,y = pct_qual_without_solar)) +
  geom_point() 
```

## Is there a statistically significant difference in solar adoption for income levels in the US and/or in California?

```{r}

#cleandata_meandiffs <- cleandata |> 
  #mutate(highlow_income = case_when(income_level_ind == 1 ~ "low/mod",
                                  #  income_level_ind == 2 ~ "low/mod",
                                  #  income_level_ind == 3 ~ "high",
                                  #  income_level_ind == 4 ~ "high")) |> 
  #filter(state_name == "California")

cleandata_ca_incomestatus <- cleandata |> 
  mutate(income_status = case_when(tract_med_family_income <= 143100 ~ "low/mod",
                                   tract_med_family_income > 143100 ~ "mod/high")) |> 
  filter(state_name == "California") |> 
  mutate(log_number_of_panels_total = log(number_of_panels_total)) |> 
  mutate(income_status = as.factor(income_status))

income_solar_sum_stats <- cleandata_ca_incomestatus |> 
  group_by(income_status) |> 
  summarize(mean_pctqualwosolar = mean(pct_qual_without_solar),
            stdev_pctqualwosolar = sd(pct_qual_without_solar))

income_solar_sum_stats

highincome_mean <- income_solar_sum_stats |> 
  filter(income_status == "mod/high") |> 
  select(mean_pctqualwosolar)

lowincome_mean <- income_solar_sum_stats |> 
  filter(income_status == "low/mod") |> 
  select(mean_pctqualwosolar)

solar_dif <- highincome_mean - lowincome_mean # point estimate
  
# c. standard error and test statistic

  # computing n and st dev for the north central region:
stdev_highincome <- income_solar_sum_stats |> 
  filter(income_status == "mod/high") |> 
  select(stdev_pctqualwosolar)

n_highincome <- cleandata_ca_incomestatus |> 
  filter(income_status == "mod/high") |> count()

  # computing the n and st dev for the south region:
stdev_lowincome <- income_solar_sum_stats |> 
  filter(income_status == "low/mod") |> 
  select(stdev_pctqualwosolar)

n_lowincome <- cleandata_ca_incomestatus |> 
  filter(income_status == "low/mod") |> count()
  
  # standard error:
SE <- as.numeric(sqrt(stdev_highincome^2 / n_highincome + stdev_lowincome^2/n_lowincome))

  # test statistic:
zscore <- as.numeric((solar_dif - 0) / SE)

t_test_output <- t.test(pct_qual_without_solar ~ income_status, data = cleandata_ca_incomestatus, conf.level = 0.95)
t_test_p_value <- t_test_output$p.value


```

\^ that could mean that there are way bigger buildings or energy use that leads to this gap. for california also look at difference in means of solar install COUNTS in rich vs poor census tracts

```{r}


income_solar_sum_stats <- cleandata_ca_incomestatus |> 
  group_by(income_status) |> 
  summarize(mean_mediansolarpanelcount = mean(number_of_panels_total),
            stdev_mediansolarpanelcount = sd(number_of_panels_total))

income_solar_sum_stats

highincome_mean <- income_solar_sum_stats |> 
  filter(income_status == "mod/high") |> 
  select(mean_mediansolarpanelcount)

lowincome_mean <- income_solar_sum_stats |> 
  filter(income_status == "low/mod") |> 
  select(mean_mediansolarpanelcount)

solar_dif <- highincome_mean - lowincome_mean # point estimate
  
# c. standard error and test statistic

  # computing n and st dev for the north central region:
stdev_highincome <- income_solar_sum_stats |> 
  filter(income_status == "mod/high") |> 
  select(stdev_mediansolarpanelcount)

n_highincome <- cleandata_ca_incomestatus |> 
  filter(income_status == "mod/high") |> count()

  # computing the n and st dev for the south region:
stdev_lowincome <- income_solar_sum_stats |> 
  filter(income_status == "low/mod") |> 
  select(stdev_mediansolarpanelcount)

n_lowincome <- cleandata_ca_incomestatus |> 
  filter(income_status == "low/mod") |> count()
  
  # standard error:
SE <- as.numeric(sqrt(stdev_highincome^2 / n_highincome + stdev_lowincome^2/n_lowincome))

  # test statistic:
zscore <- as.numeric((solar_dif - 0) / SE)


t_test_output <- t.test(existing_installs_count ~ income_status, data = no_outliers_n, conf.level = 0.95)
t_test_p_value <- t_test_output$p.value

# now t's run a regression on number of panels (or log of number of panels)
  # first check dist
ggplot(data = cleandata, aes(x = number_of_panels_total)) +
  geom_histogram() # essentially no left tail, large right skew, good candidate for a log transformation to get data normal
  
  # the distribution looks more normal when we log it:
ggplot(data = cleandata, aes(x = log(number_of_panels_total))) +
  geom_histogram() 

mod_ca <- lm(number_of_panels_total ~ tract_med_family_income + yearly_sunlight_kwh_median, data = cleandata_ca_incomestatus)
summary(mod_ca)

res_ca <- resid(mod_ca)

ggplot(data = mod_ca, aes(sample = res_ca)) +
  geom_qq() +
  geom_qq_line()

mod_allstates <- lm(number_of_panels_total ~ tract_med_family_income + yearly_sunlight_kwh_median + tract_med_family_income:state_code + state_code, data = cleandata)
summary(mod_allstates)
```

conclusion: there is more rooftop solar potential in rich neighborhoods and there are also more solar panels in rich neigborhoods

1.  

```{r}
cleandata_meandiffs <- cleandata |> 
  mutate(highlow_income = case_when(income_level_ind == 1 ~ "low/mod",
                                    income_level_ind == 2 ~ "low/mod",
                                    income_level_ind == 3 ~ "high",
                                    income_level_ind == 4 ~ "high")) |> 
  filter(state_code == "01")
```

# what are the states with the highest propensity for solar?

```{r}

state_solar_propensity <- solar_income_noNAorInf |> 
  group_by(state_code) |> 
  summarise(total_census_tracts = n(), total_count_qualified = sum(count_qualified)) |> 
  mutate(average_qualified = total_count_qualified/total_census_tracts) |> 
  rename(GEOID = state_code)

  # join to the US state polygons and then divide the average_qualified by state area
state_solar_propensity_spatial <- merge(us_states, state_solar_propensity) |> 
  mutate(average_qualified_pop_adjusted = (total_count_qualified/total_pop_15))

tm_shape(state_solar_propensity_spatial) +
  tm_polygons(col = "average_qualified_pop_adjusted")

state_solar_propensity_spatial
```

```{r}
library(sf)
library(dplyr)
library(purrr)
library(raster)
library(osmdata)
library(spDataLarge)

library(rgdal)
library(gdal)
library(tigris)

library(tigris)
library(leaflet)

rose_island <- blocks(state = "AS", cb = TRUE)
plot(rose_island)
leaflet(rose_island) %>%
addTiles() %>%
addPolygons()

```

# for further study:

-   google project sunroof could look at residential vs non-residental by using building size
-   limitations of my study: census tracts have so much diversity, dont know renting status

# my next step:

-   run one of those regressions where the impact of income and kwh changes based off of the state we're in? or graph the coeffecients for each state to see the impact
-   add a shapefile that has the polygons for each census tract (multipolygon shapefile of census tracts?)
-   add the shape file and then i can look at area to determine average solar capacity per state
-   look at if there is a significant difference in means between rich people and less affluent people when normalizing for the state / solar irradiance itself

```{r}


#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
no_outliers <- subset(cleandata_ca_incomestatus, cleandata_ca_incomestatus$yearly_sunlight_kwh_median > (Q1 - 1.5*IQR) & cleandata_ca_incomestatus$yearly_sunlight_kwh_median < (Q3 + 1.5*IQR))

mod_ca <- lm(pct_qual_without_solar ~ income_status + yearly_sunlight_kwh_median, data = no_outliers)
summary(mod_ca)

no_outliers %>% 
  ggplot(aes(x = yearly_sunlight_kwh_median, y = pct_qual_without_solar, color = income_status)) +
  #geom_point(alpha = 0.5) +
  geom_line(data = augment(mod_ca), aes(y = .fitted, color = income_status)) + 
  labs(x = "yearly_sunlight_kwh_median",
       y = "pct_qual_without_solar") +
  scale_colour_discrete("income_status")

res <- resid(mod_ca)
#qqnorm(res)

#plot(mod, which=2)

ggplot(data = mod_ca, aes(sample = res)) +
  geom_qq() +
  geom_qq_line()

no_outliers |> 
  ggplot(aes(x=yearly_sunlight_kwh_median, y = pct_qual_without_solar, color = income_status)) +
  geom_point() +
  geom_line(data = augment(mod_ca), aes(y=.fitted))

# number of solar panels mod

#find Q1, Q3, and interquartile range for values in column A
Q1_n <- quantile(no_outliers$number_of_panels_total, .25)
Q3_n <- quantile(no_outliers$number_of_panels_total, .75)
IQR_n <- IQR(no_outliers$number_of_panels_total)

#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
no_outliers_n <- subset(no_outliers, no_outliers$number_of_panels_total > (Q1_n - 1.5*IQR_n) & no_outliers$number_of_panels_total < (Q3_n + 1.5*IQR_n))


mod_ca_number <- lm(number_of_panels_total ~ income_status + yearly_sunlight_kwh_median, data = no_outliers_n)
summary(mod_ca_number)


no_outliers_n |> 
  ggplot(aes(x=yearly_sunlight_kwh_median, y = number_of_panels_total, color = income_status)) +
  #geom_point() +
  geom_line(data = augment(mod_ca_number), aes(y=.fitted))
```

```{r}
#find Q1, Q3, and interquartile range for values in column A
Q1 <- quantile(cleandata_ca_incomestatus$yearly_sunlight_kwh_median, .25)
Q3 <- quantile(cleandata_ca_incomestatus$yearly_sunlight_kwh_median, .75)
IQR <- IQR(cleandata_ca_incomestatus$yearly_sunlight_kwh_median)

#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
no_outliers <- subset(cleandata_ca_incomestatus, cleandata_ca_incomestatus$yearly_sunlight_kwh_median > (Q1 - 1.5*IQR) & cleandata_ca_incomestatus$yearly_sunlight_kwh_median < (Q3 + 1.5*IQR))

mod_ca <- lm(pct_qual_without_solar ~ income_status + yearly_sunlight_kwh_median, data = no_outliers)
summary(mod_ca)

mod_ca  %>%
  summary() %>% 
  xtable() %>% 
  kable()

no_outliers %>% 
  ggplot(aes(x = yearly_sunlight_kwh_median, y = pct_qual_without_solar, color = income_status)) +
  geom_point(alpha = 0.5) +
  geom_line(data = augment(mod_ca), aes(y = .fitted, color = income_status)) + 
  labs(x = "Median Rooftop Yearly Solar Generation Potential (kWh)",
       y = "Percent of Rooftops Suitable for \n Solar that Already Have Solar",
       title = "Solar Potential Actualized") +
  scale_colour_discrete(name = "Income Status",
                        labels = c("Low to Moderate", "Moderate to High")) +
  theme_classic()

res <- resid(mod_ca)
#qqnorm(res)

#plot(mod, which=2)

ggplot(data = mod_ca, aes(sample = res)) +
  geom_qq() +
  geom_qq_line()

no_outliers |> 
  ggplot(aes(x=yearly_sunlight_kwh_median, y = pct_qual_without_solar, color = income_status)) +
  geom_point() +
  geom_line(data = augment(mod_ca), aes(y=.fitted))

# number of solar panels mod

#find Q1, Q3, and interquartile range for values in column A
Q1_n <- quantile(no_outliers$existing_installs_count, .25)
Q3_n <- quantile(no_outliers$existing_installs_count, .75)
IQR_n <- IQR(no_outliers$existing_installs_count)

#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
no_outliers_n <- subset(no_outliers, no_outliers$existing_installs_count > (Q1_n - 1.5*IQR_n) & no_outliers$existing_installs_count < (Q3_n + 1.5*IQR_n))


mod_ca_number <- lm(existing_installs_count ~ income_status + yearly_sunlight_kwh_median, data = no_outliers_n)
summary(mod_ca_number)


no_outliers_n |> 
  ggplot(aes(x=yearly_sunlight_kwh_median, y = number_of_panels_total, color = income_status)) +
  #geom_point() +
  geom_line(data = augment(mod_ca_number), aes(y=.fitted))
```

ca_data_inflogrm \<- cleandata_inflogrm \|\> R

filter(state_name == "California")

ggplot(data = ca_data_inflogrm, aes(x = pct_qual_without_solar)) +

geom_histogram()

ggplot(data = ca_data_inflogrm, aes(x = log_pct_qual_without_solar)) +

geom_histogram()

ggplot(data = ca_data_inflogrm, aes(sample = log_pct_qual_without_solar)) +

geom_qq() +

geom_qq_line()

mod_ca \<- lm(log_pct_qual_without_solar \~ income_status + yearly_sunlight_kwh_median + income_status:yearly_sunlight_kwh_median, data = ca_data_inflogrm)

summary(mod_ca)

\# residuals are normally distributed

res_ca \<- resid(mod_ca)

ggplot(data = mod_ca, aes(sample = res_ca)) +

geom_qq() +

geom_qq_line()

#mod_heart \<- glm(pct_qual_without_solar \~ income_level_ind + yearly_sunlight_kwh_median, data = ca_data, family= gaussian(link='log'))

#summary(mod_heart)

summary(mod_ca)

ggplot(data = ca_data_inflogrm, aes(x = tract_med_family_income,y = log_pct_qual_without_solar, color = yearly_sunlight_kwh_median)) +

geom_point() +

geom_line(data = augment(mod_ca), aes(y = .fitted, color = yearly_sunlight_kwh_median))

ggplot(data = ca_data_inflogrm, aes(x = yearly_sunlight_kwh_median,y = pct_qual_without_solar)) +

geom_point()

\## categorical income

cleandata_inflogrm_ca \<- cleandata_ca_incomestatus

cleandata_inflogrm_ca[is.na(cleandata_inflogrm_ca) \| cleandata_inflogrm_ca == "Inf" \| cleandata_inflogrm_ca == "-Inf"] \<- NA

cleandata_inflogrm_ca \<- cleandata_inflogrm_ca \|\> drop_na()
